Challenges to the Two-Infall Scenario by Large Stellar Age Catalogs

### 2025-12-02 src/tex/ms.tex

While previous studies have explored a large parameter space, most have assumed that the Galactic disk never had a substantial outflow. This decision is based on hydrodynamic simulations of Galactic fountains produced by CC SNe, which predict that ejected material falls back onto the disk on relatively short timescales \citep{spitoni_galactic_2008,spitoni_effects_2009} close to its point of origin \citep{melioli_hydrodynamical_2008,melioli_hydrodynamical_2009}, suggesting that the effect on GCE should be minimal. However, the effects of feedback in simulations are sensitive to its implementation \citep[e.g.,][]{li_effects_2020,hu_code_2023}; other simulations of MW-like galaxies with different prescriptions do produce mass-loaded outflows \citep[e.g.,][]{brook_hierarchical_2011,gutcke_nihao_2017,nelson_first_2019,peschken_angular_2021,kopenhafer_seeking_2023}. Empirically, mass-loading is observed in nearby starburst galaxies \citep[e.g.,][]{lopez_temperature_2020,cameron_duvet_2021,lopez_x-ray_2023} but not MW-like systems because the predicted column densities are below current detection limits \citep[see reviews by][]{veilleux_cool_2020,thompson_theory_2024}. Even if the MW is not currently ejecting a substantial outflow, it may have done so in the past.

By neglecting Galactic outflows, previous studies of the two-infall scenario have been constrained in their choice of nucleosynthetic yields \citep{francois_evolution_2004} because of the yield--outflow degeneracy \citep[e.g.,][]{hartwick_chemical_1976,cooke_primordial_2022,johnson_dwarf_2023,sandford_strong_2024}. \citet{weinberg_equilibrium_2017} showed that the equilibrium metallicity is primarily set by the yields and the outflow mass-loading factor; proportionally raising or lowering both may affect the time-dependence of abundance evolution but not the end-point. This degeneracy prohibits direct estimates of the yields from GCE models unless the effect of outflows is assumed to be insignificant. The predicted yields from CCSN models can vary substantially depending on the choice of initial mass function \citep{vincenzo_modern_2016} and the physics of black hole formation \citep{griffith_impact_2021}, yet few studies have investigated the effect of varying the yield scale on two-infall scenario predictions. Varying the yield scale while maintaining an evolutionary end point consistent with observations is straightforwardly achieved with simultaneous variations in the strength of outflows.

@@

As stellar age estimation techniques have improved over recent years, large catalogs have become available with age estimates for stars across a wide swath of the Galaxy. In a challenge to the traditional view of GCE, which expects the ISM metallicity to continually increase over time, these catalogs suggest that stellar metallicities are remarkably consistent with age across most of the thin disk \citep[e.g.,][]{spina_mapping_2022,magrini_gaia-eso_2023,willett_evolution_2023,carbajo-hijarrubia_occaso_2024,gallart_chronology_2024}. \citet{johnson_milky_2024} demonstrated that these results are readily explained by a scenario in which the Galaxy reaches a state of chemical equilibrium. In this so-called ``equilibrium scenario,'' the local metallicity is driven by the ratio of star formation to accretion at a given radius, which evolves to a constant over Gyr timescales. Whether the equilibrium metallicity is regulated by outflows, as in the  \citet{johnson_milky_2024} models, or by other factors such as a radial gas flow \citep{spitoni_effects_2011,bilitewski_radial_2012,sharda_interplay_2024}, the current data suggest that metal abundances in the ISM evolved minimally during the past $\sim8\Gyr$.

@@

The availability of large stellar catalogs with [Fe/H], \aFe, {\it and} age covering a large portion of the Galactic disk enables much more comprehensive tests of the two-infall GCE scenario (and other models). Our study draws on the steadiness of the Milky Way stellar metallicity gradient found by \citet{johnson_milky_2024} and empirical constraints on the yield scale from \citet{weinberg_scale_2024}. We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the composition of the circumgalactic medium, and the local disk mass surface density ratio on GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss possible extensions to our models in Section \ref{sec:discussion}, and we and summarize our conclusions in Section \ref{sec:conclusions}.

@@

The population-averaged nucleosynthetic yields of CCSNe, $y_{\rm X}^{\rm CC}$, are uncertain to a degree that is significant for chemical evolution models. \citet{weinberg_scale_2024} used a measurement of the mean Fe yield of CC SNe by \citet{rodriguez_iron_2023} and the plateau in stellar \aFe abundances at low metallicity to infer population-averaged yields of $y/Z_\odot\approx1$\footnote{
    In this work, we will use the $y/Z_\odot$ notation to refer to the scale set by the massive star $\alpha$-element yields; i.e., $y_{\alpha}^{\rm CC}/Z_{\alpha,\odot}$. We also clarify that these yields refer to the net metal production by stellar populations; the return of previously produced metals in the envelopes of dying stars is handled separately by {\tt VICE}.
}\textemdash in other words, for every $1\Msun$ of stars formed, massive stars release a quantity of newly-synthesized $\alpha$-elements (e.g., O or Mg) equal to their mass fraction in the Sun. However, \citet{johnson_milky_2024} found that their GCE models with this yield scale approach present-day abundances too slowly to match the observed age--metallicity relation. Previous multi-zone models using {\tt VICE} \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024} adopted higher yields ($y/Z_\odot\approx2.6$) based on \citet{chieffi_explosive_2004} and \citet{limongi_nucleosynthesis_2006}; however, in order to produce a realistic evolution of [O/Fe], those studies adopted an integrated SN Ia rate which is high compared to the measurement of \citet{maoz_star_2017}.

@@

\citet{johnson_milky_2024} used variations in $\eta$ with $R_{\rm gal}$ to produce variations in $Z_{\rm O,eq}$.
To produce an exponentially declining radial [O/H] gradient, we adopt a prescription of
\begin{equation}
    \eta(R_{\rm gal}) = \eta_\odot \exp\Big(\frac{R_{\rm gal}-R_\odot}{R_\eta}\Big)
\end{equation}
for the outflow mass-loading factor,
where $R_\eta$ is the exponential outflow scale radius and $R_\odot=8\kpc$. We adopt $R_\eta=5$ kpc, a lower value than in \citet{johnson_milky_2024}, so that our $y/Z_\odot=1$ model produces a radial abundance gradient of $\nabla\mathOH_{\rm eq}\approx-0.06\,{\rm dex}\kpc^{-1}$, consistent with recent measurements from HII regions \citep{mendez-delgado_gradients_2022} and stars \citep{myers_open_2022,johnson_milky_2024}. We note, however, that mass-loaded outflows are not a necessary ingredient for the results of this study. We find that a one-zone model with the fiducial parameters, $\eta=0$, and \yZ{0.8}, predicts a similar abundance evolution and nearly identical stellar abundance distributions to the fiducial model with $\eta=0.2$ and \yZ{1}.

@@

In most of our models, we assume the infalling gas is pristine (i.e., $Z_{\rm in}=0$). However, previous GCE studies suggest that some level of enrichment of the infalling gas can improve agreement with observations \citep[e.g.,][]{palla_chemical_2020,johnson_milky_2024,spitoni_remind_2024}. The circumgalactic medium (CGM) from which the infalling gas is drawn could be previously enriched, possibly from contributions from Galactic outflows, gas stripped from dwarf galaxies, or from SNe in the halo. The Milky Way's CGM is diffuse, multiphase, and inhomogeneous, making it difficult to study \citep[e.g.,][]{tumlinson_circumgalactic_2017,mathur_probing_2022}; still, recent observations have confirmed the existence of metals at non-Solar abundance ratios in the CGM \citep[e.g.,][]{das_discovery_2019,das_hot_2021,gupta_supervirial_2021}. We investigate models where the infalling gas is pre-enriched to a metallicity
\begin{equation}
    \label{eq:pre-enrichment}
    Z_{\rm in}(t) = (1 - e^{-t/\tau_{\rm rise}}) Z_\odot 10^{\mathXH_{\rm CGM}}.
\end{equation}
In this case, the metallicity rises from 0 with a timescale $\tau_{\rm rise}=2\,{\rm Gyr}$ and plateaus at $\mathFeH_{\rm CGM}$. We assume the Solar-scaled abundance of the accreted gas is the same for all elements (i.e., $\mathXH_{\rm CGM}=\mathOH_{\rm CGM}$).

@@

\begin{deluxetable*}{lCCcDDD}
    \tablecaption{Summary of our multi-zone models and relevant parameters.\label{tab:models}}
    \tablehead{
        \colhead{Name} & \colhead{$y/Z_\odot$} & \colhead{$\eta_\odot$} & \colhead{SN Ia DTD} & \multicolumn2c{$\sigma_{\rm RM8}$ [kpc]} & \multicolumn2c{$f_\Sigma(R_\odot)$} & \multicolumn2c{${\rm [X/H]}_{\rm CGM}$}
    }
    \decimals
    \startdata
        {\tt yZ1-fiducial}  & 1 & 0.2   & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & $-\infty$   \\
        {\tt yZ1-migration} & 1 & 0.2   & Equation \ref{eq:plateau-dtd}  & 5.0   & 0.12  & $-\infty$   \\
        {\tt yZ1-diskratio} & 1 & 0.2   & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.5   & $-\infty$   \\
        {\tt yZ1-preenrich} & 1 & 0.6   & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & -0.5      \\
        {\tt yZ1-best}      & 1 & 0.4   & Equation \ref{eq:plateau-dtd}  & 3.6   & 0.25  & -0.7      \\
        \hline
        {\tt yZ2-fiducial}  & 2 & 1.4   & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & $-\infty$   \\
        {\tt yZ2-powerlaw}  & 2 & 1.4   & Equation \ref{eq:powerlaw-dtd} & 2.68  & 0.12  & $-\infty$   \\
        {\tt yZ2-diskratio} & 2 & 1.4   & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.5   & $-\infty$   \\
        {\tt yZ2-preenrich} & 2 & 2.4   & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & -0.5      \\
        {\tt yZ2-best}      & 2 & 1.8   & Equation \ref{eq:plateau-dtd}  & 3.6   & 0.25  & -0.7      \\
        \hline
        {\tt yZ3-fiducial}  & 3 & 2.4   & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & $-\infty$   \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

Figure \ref{fig:yield-outflow} shows the mode of the APOGEE [O/H], [Fe/H], and [O/Fe] distributions as a function of age at the Solar annulus, using the NN ages from \citet{leung_variational_2023}. As shown by \citet{johnson_milky_2024}, the peak of the MDF is less susceptible to modification by radial migration, making it a more reliable proxy for ISM chemistry at a given lookback time than the mean or median.
We over-plot the ISM abundance evolution in our $y/Z_\odot=1,2,$ and $3$ models with the fiducial parameters at $R_{\rm gal}=8\kpc$ for comparison. The models are in broad agreement with each other and with the data at lookback times of $\lesssim5\Gyr$. At ages of $\sim5-8\Gyr$, the models underpredict the metallicities of the observed stars by up to half an order of magnitude, a natural consequence of the dilution associated with the second infall event. This discrepancy is modestly reduced in the higher $y/Z_\odot$ models, which re-enrich more rapidly. However, these models exacerbate tensions with the age--[O/Fe] relation because the O abundance increases more before enrichment from second-infall SNe Ia becomes important, making the $\alpha$-enhancement from the ensuing starburst more pronounced. Overall, none of our models provide a good match to these age trends. Figure \ref{fig:yield-outflow} illustrates a point that will arise as a theme throughout the remainder of this paper, which is that the signatures of a substantial dilution event and subsequent re-enrichment, characteristic of the two-infall scenario, are simply not present in the observed age--abundance trends.

@@

Chemical evolution models that assume the \yZ{1} (empirical) yield scale already struggle to match the local age--metallicity relation even with a smooth SFH \citep[see also][]{johnson_milky_2024}. The problem is exacerbated in the two-infall case because of the delayed dilution event---the approach to equilibrium is ``reset'' by the second infall. The dilution of the ISM then gets baked into the stellar abundance record.
Figure \ref{fig:abundance-evolution-params} shows stellar age--abundance relations predicted by models with \yZ{1}. The model with the fiducial parameters (column (a); {\tt yZ1-fiducial}) shows similar discrepancies with the \citet{leung_variational_2023} NN age--[O/H] relation as in Figure \ref{fig:yield-outflow}. The evolution of [Fe/H] is similar, but the approach to the final metallicity is slower because of the additional delay imposed by SNe Ia. We smoothed simulated data points based on the median age and abundance errors in Table \ref{tab:uncertainties} to incorporate these errors in the model. The black dashed curve shows the evolution of gas phase abundances at $R_{\rm gal}=8\kpc$. Despite the effects of radial migration and statistical errors, the stellar populations mostly scatter around this gas phase curve, at least for $\tau\lesssim9\Gyr$.

@@

In contrast, the APOGEE data show remarkably little evolution in [Fe/H] up to ages of $\sim8\Gyr$ at all radii. Row (c) of Figure \ref{fig:mdf-evolution} shows that the MDF broadens with age, but its peak remains constant across this range. The mode [Fe/H] for the youngest stars (indicated by the vertical blue dotted line) is nearly the same as for the $6-8\Gyr$ old stars. At $R_{\rm gal}<7\kpc$, the MDF skews to lower [Fe/H] more noticeably with age, but its mode does not shift by more than $\sim0.1\dex$. It is difficult to draw conclusions about the outer Galaxy because the mode [Fe/H] is close to the metallicity cut at $\mathFeH>-0.4$ for luminous giants (represented by the vertical gray dashed line), which comprise the majority of stars in the sample at that distance. The remarkable consistency of the MDF over time, the result that motivated the equilibrium scenario proposed by \citet{johnson_milky_2024}, contrasts with the predictions of our fiducial models.

@@

In our models, the outflows regulate the overall metallicity in a radially dependent manner, which \citet{johnson_milky_2024} demonstrated can improve agreement with age trends across the Galactic disk. The outflow is simpler in implementation than a radial gas flow; we can simply adjust the mass loading factors up or down in combination with a given scale of yields. This approach allows us to easily consider models that predict metallicity to evolve on different timescales but reach similar abundances at the present day (see, e.g., Figure \ref{fig:yield-outflow}). We have therefore focused on models with outflows in this paper for the sake of simplicity, but constant radial gas flow models should have similar effects.

@@

We have not found a GCE model, two-infall or otherwise, that is capable of reproducing all of the observations. The ``inside-out'' models of \citet{johnson_stellar_2021} are better able to predict a slowly evolving metallicity gradient, but they do not predict an $\alpha$-bimodality. Even a smoothly-evolving SFH struggles to reproduce the flat age--metallicity relation at the lower empirical yield scale \citep{johnson_milky_2024}. We emphasize, though, that by its very construction, the two-infall scenario trades success at reproducing certain observations for failure at others.

The apparent age-independence of stellar abundances in the disk places considerable restrictions upon the two-infall parameter space because it predicts a substantial dilution event at the start of the thin disk epoch. If the equilibrium scenario of \citet{johnson_milky_2024} is accurate, then it restricts the two-infall scenario more than other GCE models. Restrictions on the two-infall scenario also translate more broadly to other merger-dominated SFHs, which must explain the MW's $\alpha$-bimodality without significantly diluting the metallicity of the ISM.

>>>>

% While previous studies have explored a large parameter space, most have assumed that the Galactic disk never had a substantial outflow. This decision is based on hydrodynamic simulations of Galactic fountains produced by CC SNe, which predict that ejected material falls back onto the disk on relatively short timescales \citep{spitoni_galactic_2008,spitoni_effects_2009} close to its point of origin \citep{melioli_hydrodynamical_2008,melioli_hydrodynamical_2009}, suggesting that the effect on GCE should be minimal. However, the effects of feedback in simulations are sensitive to its implementation \citep[e.g.,][]{li_effects_2020,hu_code_2023}; other simulations of MW-like galaxies with different prescriptions do produce mass-loaded outflows \citep[e.g.,][]{brook_hierarchical_2011,gutcke_nihao_2017,nelson_first_2019,peschken_angular_2021,kopenhafer_seeking_2023}. Empirically, mass-loading is observed in nearby starburst galaxies \citep[e.g.,][]{lopez_temperature_2020,cameron_duvet_2021,lopez_x-ray_2023} but not MW-like systems because the predicted column densities are below current detection limits \citep[see reviews by][]{veilleux_cool_2020,thompson_theory_2024}. Even if the MW is not currently ejecting a substantial outflow, it may have done so in the past.

\added{While previous studies have explored a large parameter space, most have assumed that the Galactic disk never had a substantial outflow. The presence of mass-loaded outflows in MW-like galaxies is heavily debated \citep[see discussion in][]{johnson_milky_2025}, but even if the MW is not currently ejecting a substantial outflow, it may have done so in the past.} 
By neglecting Galactic outflows, previous studies of the two-infall scenario have been constrained in their choice of nucleosynthetic yields \citep{francois_evolution_2004} because of the yield--outflow degeneracy \citep[e.g.,][]{hartwick_chemical_1976,weinberg_equilibrium_2017,cooke_primordial_2022,johnson_dwarf_2023,sandford_strong_2024}. 
% \citet{weinberg_equilibrium_2017} showed that the equilibrium metallicity is primarily set by the yields and the outflow mass-loading factor; proportionally raising or lowering both may affect the time-dependence of abundance evolution but not the end-point. 
This degeneracy prohibits direct estimates of the yields from GCE models unless the effect of outflows is assumed to be insignificant. The predicted yields from CCSN models can vary substantially depending on the choice of initial mass function \citep{vincenzo_modern_2016} and the physics of black hole formation \citep{griffith_impact_2021}, yet few studies have investigated the effect of varying the yield scale on two-infall scenario predictions. Varying the yield scale while maintaining an evolutionary end point consistent with observations is straightforwardly achieved with simultaneous variations in the strength of outflows.

@@

As stellar age estimation techniques have improved over recent years, large catalogs have become available with age estimates for stars across a wide swath of the Galaxy. In a challenge to the traditional view of GCE, which expects the ISM metallicity to continually increase over time, these catalogs suggest that stellar metallicities are remarkably consistent with age across most of the thin disk \citep[e.g.,][]{spina_mapping_2022,magrini_gaia-eso_2023,willett_evolution_2023,carbajo-hijarrubia_occaso_2024,gallart_chronology_2024}. \citet{johnson_milky_2025} demonstrated that these results are readily explained by a scenario in which the Galaxy reaches a state of chemical equilibrium. In this so-called ``equilibrium scenario,'' the local metallicity is driven by the ratio of star formation to accretion at a given radius, which evolves to a constant over Gyr timescales. Whether the equilibrium metallicity is regulated by outflows, as in the  \citet{johnson_milky_2025} models, or by other factors such as a radial gas flow \citep{spitoni_effects_2011,bilitewski_radial_2012,sharda_interplay_2024}, the current data suggest that metal abundances in the ISM evolved minimally during the past $\sim8\Gyr$.

@@

The availability of large stellar catalogs with [Fe/H], \aFe, {\it and} age covering a large portion of the Galactic disk enables much more comprehensive tests of the two-infall GCE scenario (and other models). Our study draws on the steadiness of the Milky Way stellar metallicity gradient found by \citet{johnson_milky_2025} and empirical constraints on the yield scale from \citet{weinberg_scale_2024}. We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the composition of the circumgalactic medium, and the local disk mass surface density ratio on GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss possible extensions to our models in Section \ref{sec:discussion}, and we and summarize our conclusions in Section \ref{sec:conclusions}.

@@

The population-averaged nucleosynthetic yields of CCSNe, $y_{\rm X}^{\rm CC}$, are uncertain to a degree that is significant for chemical evolution models. \citet{weinberg_scale_2024} used a measurement of the mean Fe yield of CC SNe by \citet{rodriguez_iron_2023} and the plateau in stellar \aFe abundances at low metallicity to infer population-averaged yields of $y/Z_\odot\approx1$\footnote{
    In this work, we will use the $y/Z_\odot$ notation to refer to the scale set by the massive star $\alpha$-element yields; i.e., $y_{\alpha}^{\rm CC}/Z_{\alpha,\odot}$. We also clarify that these yields refer to the net metal production by stellar populations; the return of previously produced metals in the envelopes of dying stars is handled separately by {\tt VICE}.
}\textemdash in other words, for every $1\Msun$ of stars formed, massive stars release a quantity of newly-synthesized $\alpha$-elements (e.g., O or Mg) equal to their mass fraction in the Sun. However, \citet{johnson_milky_2025} found that their GCE models with this yield scale approach present-day abundances too slowly to match the observed age--metallicity relation. Previous multi-zone models using {\tt VICE} \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024} adopted higher yields ($y/Z_\odot\approx2.6$) based on \citet{chieffi_explosive_2004} and \citet{limongi_nucleosynthesis_2006}; however, in order to produce a realistic evolution of [O/Fe], those studies adopted an integrated SN Ia rate which is high compared to the measurement of \citet{maoz_star_2017}.

@@

\citet{johnson_milky_2025} used variations in $\eta$ with $R_{\rm gal}$ to produce variations in $Z_{\rm O,eq}$.
To produce an exponentially declining radial [O/H] gradient, we adopt a prescription of
\begin{equation}
    \eta(R_{\rm gal}) = \eta_\odot \exp\Big(\frac{R_{\rm gal}-R_\odot}{R_\eta}\Big)
\end{equation}
for the outflow mass-loading factor,
where $R_\eta$ is the exponential outflow scale radius and $R_\odot=8\kpc$. We adopt $R_\eta=5$ kpc, a lower value than in \citet{johnson_milky_2025}, so that our $y/Z_\odot=1$ model produces a radial abundance gradient of $\nabla\mathOH_{\rm eq}\approx-0.06\,{\rm dex}\kpc^{-1}$, consistent with recent measurements from HII regions \citep{mendez-delgado_gradients_2022} and stars \citep{myers_open_2022,johnson_milky_2025}. We note, however, that mass-loaded outflows are not a necessary ingredient for the results of this study. We find that a one-zone model with the fiducial parameters, $\eta=0$, and \yZ{0.8}, predicts a similar abundance evolution and nearly identical stellar abundance distributions to the fiducial model with $\eta=0.2$ and \yZ{1}.

@@

In most of our models, we assume the infalling gas is pristine (i.e., $Z_{\rm in}=0$). However, previous GCE studies suggest that some level of enrichment of the infalling gas can improve agreement with observations \citep[e.g.,][]{palla_chemical_2020,johnson_milky_2025,spitoni_remind_2024}. The circumgalactic medium (CGM) from which the infalling gas is drawn could be previously enriched, possibly from contributions from Galactic outflows, gas stripped from dwarf galaxies, or from SNe in the halo. The Milky Way's CGM is diffuse, multiphase, and inhomogeneous, making it difficult to study \citep[e.g.,][]{tumlinson_circumgalactic_2017,mathur_probing_2022}; still, recent observations have confirmed the existence of metals at non-Solar abundance ratios in the CGM \citep[e.g.,][]{das_discovery_2019,das_hot_2021,gupta_supervirial_2021}. We investigate models where the infalling gas is pre-enriched to a metallicity
\begin{equation}
    \label{eq:pre-enrichment}
    Z_{\rm in}(t) = (1 - e^{-t/\tau_{\rm rise}}) Z_\odot 10^{\mathXH_{\rm CGM}}.
\end{equation}
In this case, the metallicity rises from 0 with a timescale $\tau_{\rm rise}=2\,{\rm Gyr}$ and plateaus at $\mathFeH_{\rm CGM}$. We assume the Solar-scaled abundance of the accreted gas is the same for all elements (i.e., $\mathXH_{\rm CGM}=\mathOH_{\rm CGM}$).

@@

\begin{deluxetable*}{lCCCcDDD}
    \tablecaption{Summary of our multi-zone models and relevant parameters.\label{tab:models}}
    \tablehead{
        \colhead{Name} & \colhead{$y/Z_\odot$} & \colhead{$\eta_\odot$} & \colhead{$t_{\rm max}$ [Gyr]} & \colhead{SN Ia DTD} & \multicolumn2c{$\sigma_{\rm RM8}$ [kpc]} & \multicolumn2c{$f_\Sigma(R_\odot)$} & \multicolumn2c{${\rm [X/H]}_{\rm CGM}$}
    }
    \decimals
    \startdata
        {\tt yZ1-fiducial}  & 1 & 0.2   & 4.0 & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & $-\infty$   \\
        {\tt yZ1-migration} & 1 & 0.2   & 4.0 & Equation \ref{eq:plateau-dtd}  & 5.0   & 0.12  & $-\infty$   \\
        {\tt yZ1-diskratio} & 1 & 0.2   & 4.0 & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.5   & $-\infty$   \\
        {\tt yZ1-preenrich} & 1 & 0.6   & 4.0 & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & -0.5      \\
        {\tt yZ1-best}      & 1 & 0.4   & 4.0 & Equation \ref{eq:plateau-dtd}  & 3.6   & 0.25  & -0.7      \\
        \hline
        {\tt yZ2-fiducial}  & 2 & 1.4   & 4.0 & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & $-\infty$   \\
        {\tt yZ2-earlyonset}& 2 & 1.4   & 3.0 & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & $-\infty$   \\
        {\tt yZ2-powerlaw}  & 2 & 1.4   & 3.0 & Equation \ref{eq:powerlaw-dtd} & 2.68  & 0.12  & $-\infty$   \\
        {\tt yZ2-diskratio} & 2 & 1.4   & 3.0 & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.5   & $-\infty$   \\
        {\tt yZ2-preenrich} & 2 & 2.4   & 3.0 & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & -0.5      \\
        {\tt yZ2-best}      & 2 & 1.8   & 3.0 & Equation \ref{eq:plateau-dtd}  & 3.6   & 0.25  & -0.7      \\
        \hline
        {\tt yZ3-fiducial}  & 3 & 2.4   & 4.0 & Equation \ref{eq:plateau-dtd}  & 2.68  & 0.12  & $-\infty$   \\
    \enddata
\end{deluxetable*}
\vspace{-24pt}

@@

Figure \ref{fig:yield-outflow} shows the mode of the APOGEE [O/H], [Fe/H], and [O/Fe] distributions as a function of age at the Solar annulus, using the NN ages from \citet{leung_variational_2023}. As shown by \citet{johnson_milky_2025}, the peak of the MDF is less susceptible to modification by radial migration, making it a more reliable proxy for ISM chemistry at a given lookback time than the mean or median.
We over-plot the ISM abundance evolution in our $y/Z_\odot=1,2,$ and $3$ models with the fiducial parameters at $R_{\rm gal}=8\kpc$ for comparison. The models are in broad agreement with each other and with the data at lookback times of $\lesssim5\Gyr$. At ages of $\sim5-8\Gyr$, the models underpredict the metallicities of the observed stars by up to half an order of magnitude, a natural consequence of the dilution associated with the second infall event. This discrepancy is modestly reduced in the higher $y/Z_\odot$ models, which re-enrich more rapidly. However, these models exacerbate tensions with the age--[O/Fe] relation because the O abundance increases more before enrichment from second-infall SNe Ia becomes important, making the $\alpha$-enhancement from the ensuing starburst more pronounced. Overall, none of our models provide a good match to these age trends. Figure \ref{fig:yield-outflow} illustrates a point that will arise as a theme throughout the remainder of this paper, which is that the signatures of a substantial dilution event and subsequent re-enrichment, characteristic of the two-infall scenario, are simply not present in the observed age--abundance trends.

@@

Chemical evolution models that assume the \yZ{1} (empirical) yield scale already struggle to match the local age--metallicity relation even with a smooth SFH \citep[see also][]{johnson_milky_2025}. The problem is exacerbated in the two-infall case because of the delayed dilution event---the approach to equilibrium is ``reset'' by the second infall. The dilution of the ISM then gets baked into the stellar abundance record.
Figure \ref{fig:abundance-evolution-params} shows stellar age--abundance relations predicted by models with \yZ{1}. The model with the fiducial parameters (column (a); {\tt yZ1-fiducial}) shows similar discrepancies with the \citet{leung_variational_2023} NN age--[O/H] relation as in Figure \ref{fig:yield-outflow}. The evolution of [Fe/H] is similar, but the approach to the final metallicity is slower because of the additional delay imposed by SNe Ia. We smoothed simulated data points based on the median age and abundance errors in Table \ref{tab:uncertainties} to incorporate these errors in the model. The black dashed curve shows the evolution of gas phase abundances at $R_{\rm gal}=8\kpc$. Despite the effects of radial migration and statistical errors, the stellar populations mostly scatter around this gas phase curve, at least for $\tau\lesssim9\Gyr$.

@@

In contrast, the APOGEE data show remarkably little evolution in [Fe/H] up to ages of $\sim8\Gyr$ at all radii. Row (c) of Figure \ref{fig:mdf-evolution} shows that the MDF broadens with age, but its peak remains constant across this range. The mode [Fe/H] for the youngest stars (indicated by the vertical blue dotted line) is nearly the same as for the $6-8\Gyr$ old stars. At $R_{\rm gal}<7\kpc$, the MDF skews to lower [Fe/H] more noticeably with age, but its mode does not shift by more than $\sim0.1\dex$. It is difficult to draw conclusions about the outer Galaxy because the mode [Fe/H] is close to the metallicity cut at $\mathFeH>-0.4$ for luminous giants (represented by the vertical gray dashed line), which comprise the majority of stars in the sample at that distance. The remarkable consistency of the MDF over time, the result that motivated the equilibrium scenario proposed by \citet{johnson_milky_2025}, contrasts with the predictions of our fiducial models.

@@

In our models, the outflows regulate the overall metallicity in a radially dependent manner, which \citet{johnson_milky_2025} demonstrated can improve agreement with age trends across the Galactic disk. The outflow is simpler in implementation than a radial gas flow; we can simply adjust the mass loading factors up or down in combination with a given scale of yields. This approach allows us to easily consider models that predict metallicity to evolve on different timescales but reach similar abundances at the present day (see, e.g., Figure \ref{fig:yield-outflow}). We have therefore focused on models with outflows in this paper for the sake of simplicity, but constant radial gas flow models should have similar effects.

@@

We have not found a GCE model, two-infall or otherwise, that is capable of reproducing all of the observations. The ``inside-out'' models of \citet{johnson_stellar_2021} are better able to predict a slowly evolving metallicity gradient, but they do not predict an $\alpha$-bimodality. Even a smoothly-evolving SFH struggles to reproduce the flat age--metallicity relation at the lower empirical yield scale \citep{johnson_milky_2025}. We emphasize, though, that by its very construction, the two-infall scenario trades success at reproducing certain observations for failure at others.

The apparent age-independence of stellar abundances in the disk places considerable restrictions upon the two-infall parameter space because it predicts a substantial dilution event at the start of the thin disk epoch. If the equilibrium scenario of \citet{johnson_milky_2025} is accurate, then it restricts the two-infall scenario more than other GCE models. Restrictions on the two-infall scenario also translate more broadly to other merger-dominated SFHs, which must explain the MW's $\alpha$-bimodality without significantly diluting the metallicity of the ISM.
